Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-08-26 2024/OPTIMA/OPTIMA_2024_MCO.tex
submitted

%\begin{equation}
%\label{eq:06}
%F(y) = \{ F(\lambda_1,y),F(\lambda_2,y), \dots ,F(\lambda_\tau,y)\}
%\end{equation}

@@

\textit{Remark 2.} It follows from relationship between the Lipschitz and H{\" o}lder constants of the functions $\varphi(y)$ and $\varphi(y(x))$, as well as from the theorem of convergence of the global search algorithm (see \cite{Strongin2000}), that parameter $r$ from (\ref{eq:10}) should satisfy the necessary condition $r>2^{2-1/N}$. 
In the test experiments described in Section~\ref{sec:5} we used the value $r=3$, which ensured the correct solution of the entire series of problems. In applied optimization problems the value $r=3$ also ensures the solution in the most cases.

%Theoretical justification of this algorithm is provided in \cite{Strongin2000}.

@@

As part of the proposed approach, the machine learning model is built according to the rules similar to the construction of the SVC model from section \ref{ssec:41}. As before, all trial points are divided into two classes -- points of the Pareto set and points that do not belong to the Pareto set. The classifier is trained on the selected classes. Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class. Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points
%$y_i=y(x_i)$ and $y_{i-1}=y(x_{i-1})$
corresponding to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.

A significant advantage of this approach is the possibility to easily replace both the machine learning model itself and its learning rules. In the framework of the study, several variants of building models that potentially improve the factor $R_{ps}(i)$ to beter solve the MCO problems.

@@

%Computations were performed with the support of the scikit-learn 0.24.2 framework.

%of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.  and the optimization software iOpt \cite{iOpt_url}.

@@

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage{graphicx}
\begin{table}[t]
\centering
\caption{Results of computational experiments}
\label{tab:1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{cccccc}
\hline
\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV index} & \textbf{Avg. iter.} & \textbf{Speedup} \\ \hline
ML\_MGSA                                                                   & -          & -    & 114.91  & 2269.82   & 1.00           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of fixed weights}}                                                                \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Fixed weights)}                            & Linear     & 0.04 & 114.72  & 963.05    & 2.36           \\
                                                                           & Poly       & 0.07 & 114.73  & 871.82    & 2.60           \\
                                                                           & RBF        & 0.03 & 114.80  & 883.04    & 2.57           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights}}                                                             \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Adj. weights)}                             & Linear     & 0.04 & 114.76  & 1018.03    & 2.23           \\
                                                                           & Poly       & 0.08 & 114.73  & 844.66    & 2.69           \\
                                                                           & RBF        & 0.03 & 114.76  & 796.74    & 2.85           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights and logarithmic probabilities}}                                               \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                 & Linear     & 0.05 & 114.72  & 711.44    & 3.19           \\
                                                                           & Poly       & 0.04 & 114.79  & 854.82    & 2.66           \\
                                                                           & RBF        & 0.04 & 114.74  & 704.99    & 3.22           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments with 4-criterial MCO problems}}                                                                \\ \hline
ML\_MGSA                                                                   & -          & -    & 7819.86 & 6046.4    & 1.00           \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 7751.43 & 1925.6    & 3.14           \\
\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                 & Linear     & 0.08 & 7729.71 & 1760.5    & 3.43           \\
                                                                           & Poly       & 0.08 & 7796.20 & 4121.7    & 1.47           \\
                                                                           & RBF        & 0.08 & 7786.48 & 1364.6    & \textbf{4.43}  \\ \hline
\end{tabular}%
}
\end{table}

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
%
%\begin{thebibliography}{8}
%\bibitem{ref_article1}
%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
%
%\bibitem{ref_lncs1}
%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
%
%\bibitem{ref_book1}
%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
%Location (1999)
%
%\bibitem{ref_proc1}
%Author, A.-B.: Contribution title. In: 9th International Proceedings
%on Proceedings, pp. 1--2. Publisher, Location (2010)
%
%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
%\end{thebibliography}
\end{document}

>>>>

@@

\textit{Remark 2.} It follows from relationship between the Lipschitz and H{\" o}lder constants of the functions $\varphi(y)$ and $\varphi(y(x))$, as well as from the theorem of convergence of the global search algorithm (see \cite{Strongin2000}), that parameter $r$ from (\ref{eq:10}) should satisfy the necessary condition $r>2^{2-1/N}$. 
In the test experiments described in Section~\ref{sec:5} we used the value $r=3$, which ensured the correct solution of the entire series of problems. In applied optimization problems the value $r=3$ also provides the correct solution in most cases.

@@

As part of the proposed approach, the machine learning model is built according to the rules similar to the construction of the SVC model from section \ref{ssec:41}. As before, all trial points are divided into two classes -- points of the Pareto set and points that do not belong to the Pareto set. The classifier is trained on the selected classes. Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class. Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points
corresponding to boundaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.

A significant advantage of this approach is the possibility to easily replace both the machine learning model itself and its learning rules. In the framework of the study, several variants of building models that potentially improve the factor $R_{ps}(i)$ to better solve the MCO problems.

@@

@@

\begin{table}[t]
\centering
\caption{Results of computational experiments}
\label{tab:1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{cccccc}
\hline
\textbf{}                                                                  & \textbf{Model Kernel} & \textbf{$\alpha$} & \textbf{Avg. HV index} & \textbf{Avg. iter.} & \textbf{Speedup} \\ \hline
ML\_MGSA                                                                   & -          & -    & 114.91  & 2269.82   & 1.00           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of fixed weights}}                                                                \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Fixed weights)}                            & Linear     & 0.04 & 114.72  & 963.05    & 2.36           \\
                                                                           & Poly       & 0.07 & 114.73  & 871.82    & 2.60           \\
                                                                           & RBF        & 0.03 & 114.80  & 883.04    & 2.57           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights}}                                                             \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Adj. weights)}                             & Linear     & 0.04 & 114.76  & 1018.03    & 2.23           \\
                                                                           & Poly       & 0.08 & 114.73  & 844.66    & 2.69           \\
                                                                           & RBF        & 0.03 & 114.76  & 796.74    & 2.85           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments for the case of adjusted weights and logarithmic probabilities}}                                               \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 114.67  & 691.14    & \textbf{3.28}  \\
\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                 & Linear     & 0.05 & 114.72  & 711.44    & 3.19           \\
                                                                           & Poly       & 0.04 & 114.79  & 854.82    & 2.66           \\
                                                                           & RBF        & 0.04 & 114.74  & 704.99    & 3.22           \\ \hline
\multicolumn{6}{c}{\textbf{Experiments with 4-criterial MCO problems}}                                                                \\ \hline
ML\_MGSA                                                                   & -          & -    & 7819.86 & 6046.4    & 1.00           \\ \hline
ML\_MGSA\_Dist                                                             & Linear     & 0.01 & 7751.43 & 1925.6    & 3.14           \\
\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}                                 & Linear     & 0.08 & 7729.71 & 1760.5    & 3.43           \\
                                                                           & Poly       & 0.08 & 7796.20 & 4121.7    & 1.47           \\
                                                                           & RBF        & 0.08 & 7786.48 & 1364.6    & \textbf{4.43}  \\ \hline
\end{tabular}%
}
\end{table}

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{bibliography}
%
\end{document}
